A Fast kNN Algorithm Using Multiple Space-Filling Curves

### 2022-03-20 2022/Sensors/sensors.tex
add two references

Currently, machine learning (ML) methods are being used successfully to solve a wide range of problems in various application areas. One example of a class of problems where ML has demonstrated its effectiveness are the tasks of identifying the main properties of the phenomena, which are characterized by their stochastic nature or the presence of hidden parameters \cite{Golovenkin2020, Gonoskov2019, Kastalskiy2021}.

In many cases, solving application problems comes down to the problem of classification, i.e. assigning the object under study to one of the available classes. One of the well-known classification methods is the kNN (k-nearest neighbourhood) method. Although the earliest papers related to this method appeared more than $50$ years ago \cite{Cover1967, Cover1968}, theoretical studies of various aspects of using kNN are still going on [Gorban, Pestov]. This is not to mention hundreds of publications in which this method is used to solve applied classification problems. Over many years of use, kNN has established itself as a simple and reliable method that yields acceptable results when solving many different problems.

Among the undoubted merits of kNN is its ``explainability'', since the decision about whether a test object belongs to one class or another is clearly explained by similar properties of the test object and its nearest (in some metric) neighbours.
Along with the study of the method's theoretical properties, the issue of its effective implementation is also investigated. Thus, direct implementation of the method by using the  ``brute-force'' approach has computational complexity $O(n)$, where $n$ is the number of objects. Fast algorithms, which have computational complexity $O(log(n))$, use complex data structures of kd-tree type [kdtree-1,2] \cite{Bentley1975MultidimensionalBS}.

@@

By using multiple mappings $y^l(x), \; 0 \leq l \leq L$, the following relationship between neighborhoods on one-dimensional scales and neighborhoods in the original multidimensional domain is defined.

\textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let

\begin{equation}
|y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
\end{equation} 
where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different. Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that

\begin{equation}
 y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
\max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
\end{equation}

@@

For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned

\begin{equation}
\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \}
\end{equation} 
where $l, \; 0 \leq l \leq L$, is the number of the one-dimensional interval.

The degree of proximity of a test object to one of the known objects can be expressed in terms of the distance on the $l$-th scale.

\begin{equation}
d_i^n = |x_i^n - x^{*n}|
\end{equation}

Thus, the distance $d(\omega_i,\omega_j)$ between two objects $\omega_i$ and $\omega_j$ on a one-dimensional scale will correspond to the minimum distance among all the mappings. If we find the minimum distance among all these distances, we get the nearest neighbour

\begin{equation}
d_{min} = \min \{ d_i^n \} = \min \{ |x_i^n - x^{*n}| \}.
\end{equation}

@@

The input data of the algorithm also includes the feature vector $y^*$ corresponding to the test object $\omega^*$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{Conceptualization and methodology, K.B.; software and validation, A.S., A.Sh.; formal analysis, K.B.; investigation, A.S.; data curation, A.Sh.; writing --- original draft preparation, K.B.; writing- -- review and editing, K.B.; visualization, A.S.; funding acquisition, K.B. All authors have read and agreed to the published version of the manuscript.}

>>>>

Currently, machine learning (ML) methods are being used successfully to solve a wide range of problems in various application areas. One example of a class of problems where ML has demonstrated its effectiveness are the tasks of identifying the main properties of the phenomena, which are characterized by their stochastic nature or the presence of hidden parameters \cite{Golovenkin2020,Gonoskov2019,Kastalskiy2021}.

In many cases, solving application problems comes down to the problem of classification, i.e. assigning the object under study to one of the available classes. One of the well-known classification methods is the kNN (k-nearest neighbourhood) method. Although the earliest papers related to this method appeared more than $50$ years ago \cite{Cover1967, Cover1968}, theoretical studies of various aspects of using kNN are still going on \cite{Pestov2013,Mirkes2020}. This is not to mention hundreds of publications in which this method is used to solve applied classification problems. Over many years of use, kNN has established itself as a simple and reliable method that yields acceptable results when solving many different problems.

Among the undoubted merits of kNN is its ``explainability'', since the decision about whether a test object belongs to one class or another is clearly explained by similar properties of the test object and its nearest (in some metric) neighbours.
Along with the study of the method's theoretical properties, the issue of its effective implementation is also investigated. Thus, direct implementation of the method by using the  ``brute-force'' approach has computational complexity $O(n)$, where $n$ is the number of objects. Fast algorithms, which have computational complexity $O(log(n))$, use complex data structures of kd-tree type \cite{Bentley1975MultidimensionalBS}.

@@

By using multiple mappings $y^l(x), \; 0 \leq l \leq L$, the following relationship between neighbourhoods on one-dimensional scales and neighbourhoods in the original multidimensional domain is defined.

\textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
\begin{equation}
|y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
\end{equation} 
where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different. Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that
\begin{equation}
 y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
\max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
\end{equation}

@@

For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned 
\begin{equation}
\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \}
\end{equation} 
where $l, \; 0 \leq l \leq L$, is the number of the one-dimensional interval.

The degree of proximity of a test object to one of the known objects can be expressed in terms of the distance on the $l$-th scale.
\begin{equation}
d_i^n = |x_i^n - x^{*n}|
\end{equation}

Thus, the distance $d(\omega_i,\omega_j)$ between two objects $\omega_i$ and $\omega_j$ on a one-dimensional scale will correspond to the minimum distance among all the mappings. If we find the minimum distance among all these distances, we get the nearest neighbour
\begin{equation}
d_{min} = \min \{ d_i^n \} = \min \{ |x_i^n - x^{*n}| \}.
\end{equation}

@@

The input data of the algorithm also includes the feature vector $y^*$ corresponding to the test object $\omega^*$.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{Conceptualization and methodology, K.B.; software and validation, A.S., A.Sh.; formal analysis, K.B.; investigation, A.S.; data curation, A.Sh.; writing --- original draft preparation, K.B.; writing --- review and editing, K.B.; visualization, A.S.; funding acquisition, K.B. All authors have read and agreed to the published version of the manuscript.}
